---
title: "Exploring the TDoR Data"
author: "Paul Brennan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring the TDoR Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages.
```{r load_package, warning=FALSE}
anim <- require(gganimate)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(maps)
library(tdor)
```

## A set of code and graphs to illustrate the data in this package
[TRIGGER WARNING: VIOLENCE. MURDER]

The data can be explored through this website: [https://tdor.translivesmatter.info/]

## Plot reported deaths across the years
```{r deaths_over_year, warning=FALSE}
ggplot(tdor, aes(Year)) + geom_bar() +
    ggtitle("Deaths by year")
```



## Plot deaths by age

In some cases there is ambiguity over the age, so we have a minimum and a maximum. 

```{r deaths_by_age, warning=FALSE}
tdor %>% 
  filter(Age_min > 0 & Age_max > 0) %>%
  ggplot(aes(x = (Age_min + Age_max)/2)) + 
  geom_bar() +
  ggtitle("Deaths by age") +
  labs(y = "Deaths")

tdor %>% 
  filter(Age_min > 0 & Age_max > 0) %>%
  ggplot(aes(x = (Age_min + Age_max)/2)) +
  geom_histogram(binwidth = 5) +
  ggtitle("Deaths by age") +
  labs(y = "Deaths")
```

## Show top 10 countries for deaths
```{r deaths_by_country}
tdor %>%
  group_by(Country) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) -> by_country

ggplot(by_country[1:10,],
  aes(x = Country,
      y = n)) +
  theme_bw() +
  labs(y = "Deaths", x = "") +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ggtitle("Ten countries with the most reported deaths")
```

```{r world_map, warning=FALSE}
world <- ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +
  theme_map() 

map <- world +
  geom_point(aes(x = Longitude, y = Latitude),
             data = tdor, 
             colour = 'purple', alpha = .2) +
  ggtitle("Map of deaths (2006-2018)")

map
```


### Code to animate 2018 data
Animate by name 
```{r world_map_animation, eval = FALSE}
if (anim){
d18 <- filter(tdor, Year == 2018)

map_anim_name <- world +
  geom_point(aes(x = Longitude, y = Latitude),
             data = d18, 
             colour = 'red', alpha = .5) +
  labs(title = paste('{closest_state}', "died here")) +
  transition_states(
    Name,
    transition_length = 1,
    state_length = 1) + 
  shadow_mark()

# to get full animation...
animate(map_anim_name, nframes = nrow(d18)*2)

animate(map_anim_name, nframes = nrow(d18)*2) +
    anim_save(filename = 'world_map_names.gif')
}
```


### Deaths per month, as a line chart
```{r deaths_pm}
tdor %>% 
  select(Date) %>% 
  mutate(month=lubridate::floor_date(Date, unit = "1 month")) %>% 
  count(month) %>%
  right_join(
    seq.Date(from = as.Date("2006-10-01"), to=as.Date("2018-09-01"), , by="month") %>% 
  as_tibble() %>% 
  select(month=value)
  ) %>% 
  ggplot()+
  aes(x=month, y=n)+
  geom_line()
```
